---
title: "Day 7 Raster"
---

## Libraries

```{r eval=FALSE}
require(devtools)
devtools::install_github("ropensci/rerddap")
devtools::install_github("rmendels/rerddapXtracto") 
```

```{r}
library(tidyverse)
library(tidyterra)
```

## Download sea surface temperature data

I will download the optimal interpolation sea surface temperature from CoastWatch. The bounding box is off the coast of Washington, USA.

```{r}
filePath <- here::here("content", "data", "day7_sst_raster.RData")
lats <- c(40.375, 50.375)
lons <- c(-141.875, -120.875)
if (!file.exists(filePath)) {
df_info <- rerddap::info("ncdcOisst21Agg_LonPM180")
df <- rerddap::griddap("ncdcOisst21Agg_LonPM180", latitude = lats, longitude = lons, time = c("2021-06-19", "2021-06-19"), fields = "sst")$data
save(df, file=filePath)
}else{
  load(filePath)
}
```

Turn this into a raster. Need to tell raster that this is lat/long data.
```{r}
df2 <- data.frame(x=df$lon, y=df$lat, z=df$sst)
ras <- raster::rasterFromXYZ(df2, crs = "+proj=longlat")
```

```{r}
library(terra)
r <- terra::rast(ras)
names(r) <- "alt"
setMinMax(r)
slope <- terra::terrain(r, "slope", unit = "radians")
aspect <- terra::terrain(r, "aspect", unit = "radians")
hill <- terra::shade(slope, aspect, 10, 200)
names(hill) <- "shades"
```

```{r}
library(scales)

# Hillshading, but we need a palette
pal_greys <- hcl.colors(1000, "Grays")

index <- hill %>%
  mutate(index_col = rescale(shades, to = c(1, length(pal_greys)))) %>%
  mutate(index_col = round(index_col)) %>%
  pull(index_col)
# Get cols
vector_cols <- pal_greys[index]

hill_plot <- ggplot() +
  geom_spatraster(
    data = hill, fill = vector_cols, maxcell = Inf,
    alpha = 1
  )
hill_plot
```

```{r}
r_limits <- minmax(r) %>% as.vector()
# Rounded to lower and upper 500
#r_limits <- c(floor(r_limits[1] / 500), ceiling(r_limits[2] / 500)) * 500

base_plot <- hill_plot +
  geom_spatraster(data = r, maxcell = Inf) +
  scale_fill_hypso_tint_c(
    limits = r_limits,
    palette = "dem_poster",
    alpha = 0.4,
    labels = label_comma(),
    # For the legend I use custom breaks
    breaks = seq(-300, 1000, 100)
  )
base_plot
```

```{r}
library(tidyterra)
hill_plot +
  geom_spatraster(data=r, maxcell = Inf) +
  scale_fill_hypso_tint_c(limits = as.vector(minmax(r)), 
                          palette = "etopo1",
                          alpha =0.4,
                          direction = 1,
                          labels = scales::label_comma()) +
  guides(fill=guide_legend(title = "elevation", reverse = FALSE))
```
```{r}
ggplot() + 
  geom_spatraster(data=slope, maxcell = Inf) +
  scale_fill_whitebox_c(palette = "muted", alpha = 0.9, direction = 1)
  geom_spatraster(data=r, maxcell = Inf) +
  scale_fill_whitebox_c(palette = "muted", alpha = 0.9, direction = 1)
```

```{r}
hill_plot +
  geom_spatraster(data=r, maxcell = Inf) +
  scale_fill_continuous_diverging(
    palette = "Blue-Red-2", p2 = .1, alpha=.8, mid=mean(df$sst, na.rm = TRUE))
```


## Plot the raster

```{r}
library(raster)
library(ggplot2)
gg <- ggplot(df) +
  geom_raster(aes(longitude, latitude, fill = sst)) +
  scale_fill_gradient2(midpoint = mean(df$sst, na.rm = TRUE),
                       low = "blue",
                       mid = "white",
                       high = "red") +
  labs(x = NULL,
       y = NULL,
       fill = "Celcius",
       title = "Sea Surface Temperature (SST)") +
  theme_bw() +
  scale_x_continuous(limits = lons, expand = c(-0.01, -0.01)) +
  scale_y_continuous(limits = lats, expand = c(-0.01, -0.01))
gg
```

## Add a coastline

```{r}
coast <- rnaturalearth::ne_coastline(scale = 50, returnclass = "sp")
wa_or_coast <- raster::crop(coast, raster::extent(lons[1], lons[2], lats[1], lats[2]))
plot(wa_or_coast)
```

The way that ggplot2 works is to run `fortify()` on the SpatialLines object to create a data frame. Then we use `geom_path()` to plot that. But if you look at the coast, you see lots of islands. We need to add `grouping` to tell geom_path() that there are these groups of paths in the data frame.
```{r}
gg +
  geom_path(data=wa_or_coast,  aes(x=long,y=lat, grouping=id), size=1, na.rm=TRUE)
```